Convective heat transport in stratified atmospheres at low and high Mach number

### 2017-08-29 tex/nusselt_paper.tex
publication-like .tex

%\documentclass[iop]{emulateapj}
\documentclass[aps, prfluids, onecolumn, notitlepage, nofootinbib, groupedaddress, amsfonts, amssymb, amsmath]{revtex4-1}
\usepackage{graphicx}
\usepackage{bm}
\usepackage{natbib}
\usepackage{longtable}
%\usepackage[colorlinks=True, linkcolor=blue, citecolor=blue]{hyperref}
%\usepackage[all]{hypcap}

\bibliographystyle{apsrev}

@@

\newcommand\mnras{{MNRAS}}%
          % Monthly Notices of the RAS

\begin{document}
%%%%% Create nice title and abstract
\author{Evan H. Anders}
\affiliation{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affiliation{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}
\author{Benjamin P. Brown}
\affiliation{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affiliation{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}
\title{Convective heat transport in stratified atmospheres at low and high Mach number}

\begin{abstract}
We study fully compressible convection in the context of 
plane-parallel, polytropically stratified atmospheres. 
We perform a suite of 2D and 3D simulations in which we vary the initial
superadiabaticity ($\epsilon$) and the Rayleigh number (Ra) while fixing the
initial density stratification, aspect
ratio, and Prandtl number.
The evolved heat transport, 
quantified by the Nusselt number (Nu),
follows scaling relationships similar to those found in the well-studied, 
incompressible Rayleigh-B\'{e}nard problem.  This scaling holds up in both 2D and 3D and is not
appreciably affected by the magnitude of $\epsilon$. %think about last sentence.
\end{abstract}
\maketitle

%%%%% Body of the paper
\section{Introduction}
\label{sec:intro}
Convection transports energy in stellar and planetary atmospheres
where flows are compressible and
feel the atmospheric stratification.  This stratification 
is significant in regions such as
the convective envelope of the Sun, which spans 14 density scale heights.
In the bulk of these systems, particularly in the deep interior,
flows are at very low Mach number (Ma).  Unfortunately,
numerical constraints have restricted most studies of 
compressible convection to high Ma.
These prior studies \cite{graham1975, chan&all1982,
hurlburt&all1984, cattaneo&all1990, brummell&all1996,
brandenburg&all2005} have provided insight into the nature of
convection in the low temperature,
high Ma region near the Sun's surface. Few fundamental
properties of low Ma compressible convection, such as the scaling of
convective heat transport, are known.

@@

\section{Experiment} 
\label{sec:experiment}
We examine a monatomic ideal gas with an adiabatic index of
$\gamma = 5/3$ whose equation of state is $P = R\rho T$. This is consistent with the approach used in earlier work 
\cite{graham1975, chan&all1982, brandenburg&all2005,
hurlburt&all1984, cattaneo&all1990, cattaneo&all1991, brummell&all1996} 
and is the simplest stratified extension of RBC.
The atmospheres studied are initially polytropically stratified,
\begin{equation}
\begin{split}
\rho_0(z) &= \rho_{t}(1 + L_z - z)^m, \\
T_0(z)    &= T_{t}(1 + L_z - z),
\label{eqn:polytrope}
\end{split}
\end{equation}
where $m$ is the polytropic index and $L_z$ is the depth of the atmosphere.
The polytropic
index is set by the superadiabatic excess, $\epsilon = m_{ad} - m$, where
$m_{ad} = (\gamma - 1)^{-1}$ is the adiabatic value of $m$.
The height coordinate, $z$, increases upwards in the range $[0, L_z]$.
Subscript 0 indicates initial conditions and subscript $t$ indicates values
at $z = L_z$.   Stratified atmospheres have a fourth non-dimensional parameter,
the number of density scale heights, $n_{\rho} = \ln\left[\rho_0(z=0)/\rho_t\right]$.  We
specify the depth of the atmosphere, $L_z = e^{n_{\rho}/m} - 1$, by choosing
the initial value of $n_{\rho}$.
Throughout this work we set $n_{\rho} = 3$.    Satisfying hydrostatic
equilibrium sets the value of gravity, $g = RT_t (m + 1)$, which is
constant with depth.  We study atmospheres with aspect
ratios of 4 where both the $x$ and $y$ coordinates have the range $[0, 4L_z]$.
In our 2D cases, we only consider $x$ and $z$.

@@

At fixed $n_\rho$, convective dynamics are 
controlled by $\epsilon$ as well as the atmospheric diffusivities.
At a fixed value of
$\epsilon$, the diffusivities are set by the
Rayleigh number (Ra) and the Prandtl number (Pr),
\begin{equation}
\text{Ra}_{t} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu_t\chi_t},
\qquad
\text{Pr} = \frac{\nu}{\chi},
\end{equation}
where $\Delta S_0 = \epsilon\ln (1 + L_z) = \epsilon n_\rho / m$ 
is the initial specific entropy difference across the domain,
$\chi$ is the thermal diffusivity, 
and $\nu$ is the kinematic viscosity.
Throughout this work we specify
that Pr $= 1$ and is depth invariant.
The initial thermal
conductivity, $\kappa_0 = \chi \rho_0$, is
constant with depth, such that (\ref{eqn:polytrope}) is in
thermal equilibrium ($\grad\cdot[\kappa_0\grad T_0] = 0$).
By these
choices, $\nu(z) \equiv \chi(z) \equiv \chi_t / \rho_0$.
This formulation 
sets Ra at the bottom of the domain greater than
Ra$_t$ by a factor of $e^{2n_\rho}$. Henceforth
when we specify Ra we are referring to Ra$_t$.  
The full values of $\kappa = \rho\chi$ and 
$\mu = \rho\nu$ (the dynamic viscosity) change as the density 
profile evolves.  
The diffusivities scale as
$\chi_t, \nu_t \propto \sqrt{g L_z^3 (\Delta S_0 / c_P) / \text{Ra}_t}$.
Defining the thermal diffusion timescale as $t_\chi \equiv L_z^2 / \chi$, the
ratio of $t_\chi$ to the buoyancy time is
\begin{equation}
\frac{t_\chi}{t_b} =\text{Ra}_t^{1/2}.
\label{eqn:timescales}
\end{equation}
We carry out two experiments in this study. In the first,
we fix $\epsilon$ and increase Ra, thus increasing the ratio in
(\ref{eqn:timescales}). In the second, we fix Ra and vary $\epsilon$,
scaling the dynamical timescales $(t_b, t_\chi)$ as $\epsilon^{-1/2}$
relative to the speed of sound; we see this reflected in the evolved Mach
number scaling (Fig. \ref{fig:ma_v_eps}).

We use $\ln \rho$ and $T$ as our thermodynamic variables and solve the Fully 
Compressible Navier-Stokes equations,
\begin{align}
&\begin{aligned}
&\frac{\partial \ln\rho}{\partial t} + \grad\cdot\bm{u} 
    = -\bm{u}\cdot\grad\ln\rho,
	\label{eqn:continuity_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial\bm{u}}{\partial t} + \grad T - 
&\nu\grad\cdot\lilstressT - \lilstressT\cdot\grad\nu =
-\bm{u}\cdot\grad\bm{u} - T\grad\ln\rho + \bm{g} + 
\nu\lilstressT\cdot\grad\ln\rho,
\label{eqn:momentum_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial T}{\partial t} -\frac{1}{c_V}\left(\right.\chi&\left.
    \grad^2 T + \grad T\cdot\grad\chi\right) =
	-\bm{u}\cdot\grad T - (\gamma-1)T\grad\cdot{\bm{u}}
	+ \frac{1}{c_V}\left(\chi\grad T \cdot\grad\ln\rho +
	\nu\left[\lilstressT\cdot\nabla\right]\cdot\bm{u}\right), 
	\label{eqn:energy_eqn}
\end{aligned}
\end{align}
with the viscous stress tensor given by
\begin{equation}
\sigma_{ij} \equiv \left(\frac{\partial u_i}{\partial x_j} + 
\frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij}\grad\cdot\bm{u}\right).
	\label{eqn:stress_tensor}
\end{equation}
Taking an inner product of
(\ref{eqn:momentum_eqn}) with $\rho\bm{u}$ and adding it to 
$\rho c_V\times$(\ref{eqn:energy_eqn}) reveals the full energy equation,
\begin{equation}
\frac{\partial}{\partial t}\left(\rho\left[\frac{|\bm{u}|^2}{2} + c_V T + \phi\right]\right) +
\Div{\bm{F}_{\text{conv}} + \bm{F}_{\text{cond}}} = 0,
	\label{eqn:energy_eqn_full}
\end{equation}
where
$
\bm{F}_{\text{conv}} \equiv \bm{F}_{\text{enth}} + \bm{F}_{\text{KE}} + \bm{F}_{\text{PE}} + \bm{F}_{\text{visc}}
$
is the convective flux and $\bm{F}_{\text{cond}} = -\kappa \grad T$
is the conductive flux.
The individual contributions to $\bm{F}_{\text{conv}}$ are the enthalpy flux, 
$\bm{F}_{\text{enth}} \equiv \rho\bm{u}(c_V T + P/\rho)$;
the kinetic energy flux, 
$\bm{F}_{\text{KE}} \equiv \rho|\bm{u}|^2\bm{u}/2$;
the potential energy flux,
$\bm{F}_{\text{PE}} \equiv \rho\bm{u}\phi$ (with $\phi \equiv -gz$);
and the viscous flux, 
$\bm{F}_{\text{visc}} \equiv -\rho\nu\bm{u}\cdot\lilstressT$.
Understanding how each of these fluxes interact  
is crucial in characterizing convective heat transport.

We utilize the 
Dedalus\footnote{http://dedalus-project.org/} 
pseudospectral framework \cite{burns&all2016} to time-evolve  
(\ref{eqn:continuity_eqn})-(\ref{eqn:energy_eqn}) 
using an implicit-explicit (IMEX), third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
Thermodynamic variables are decomposed such that $T = T_0 + T_1$ and
$\ln\rho = (\ln\rho)_0 + (\ln\rho)_1$, 
and the velocity is $\bm{u} = w\bm{\hat{z}} + u\bm{\hat{x}} + v\bm{\hat{y}}$.
In our 2D runs, $v = 0$.
Subscript 0 variables, set by (\ref{eqn:polytrope}), 
have no time derivative and vary only in $z$.
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal, periodic) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  
Domain sizes range from
64x256 coefficients at the lowest values of 
Ra to 1024x4096 coefficients at Ra $> 10^{7}$ in 2D,
and from 64x128$^2$ to  256x512$^2$ in 3D. 
By using IMEX timestepping, we implicitly step the 
stiff linear acoustic wave contribution and are able to
efficiently study flows at high ($\sim 1$) 
and low ($\sim 10^{-4}$) Ma.  Our equations take the form
of the FC equations in \cite{lecoanet&all2014}, extended to include
$\nu$ and $\chi$ which vary with depth, and we follow the approach there.
This IMEX approach has been successfully 
tested against a nonlinear benchmark  of the compressible 
Kelvin-Helmholtz instability \cite{Lecoanet_et_al_2016_KH}.

@@

\begin{figure}[t]
\includegraphics[width=\textwidth]{./figs/ma_v_Ra.png}
\caption{The mean adiabatic Mach number of long-time-averaged profiles
is shown.  Error bars show the full range of Ma over the depth of the
atmosphere.
(a) Ma, at various values of Ra, 
is plotted as a function of $\epsilon$.
(b) Ma, at various values of $\epsilon$, 
is plotted as a function of Ra/Ra$_{\text{crit}}$.
Larger symbols with white dots designate 3D runs.
\label{fig:ma_v_eps} }
\end{figure}

@@

We measure the adiabatic Mach number (Ma = $|\bm{u}|/\sqrt{\gamma T}$),
and find that it is a strong function of 
$\epsilon$ and a weak function of Ra.  
In our 2D runs, when Ma $< 1$, we observe a scaling law of
Ma(Ra$, \epsilon) \appropto \epsilon^{1/2}$Ra$^{1/4}$.
This relation breaks down as the mean
Ma approaches 1 (Fig. \ref{fig:ma_v_eps}).  This transition
occurs near Ra/Ra$_\text{crit} \approx \{10^{2}, 10^{3}\}$ for $\epsilon = \{1, 0.5\}$.
In our limited 3D runs, Ma appears to be a function of $\epsilon$ alone, with
Ma $\appropto \epsilon^{1/2}$, so at high Ra, Ma$_{\text{3D}} < \text{Ma}_{\text{2D}}$.
We conjecture that the scaling of Ma with Ra in the 2D runs 
is due to the formation of coherent high-velocity ``spinners,'' which form
between upflow and downflow lanes in 2D.  These structures, which are reminiscent of flywheel
modes in RBC, do not appear in our 3D runs at these parameters
\cite{jones&all1976, brummell&all2002}.
Simulations in the range of Ra/Ra$_{\text{crit}} > 10^3$ at $\epsilon = 10^{-4}$
exhibited ``windy'' states of convection, in which a large-scale horizontal
shearing flow replaced the more standard upflow/downflow morphology of
convection.  Similar states have been studied in
RBC \cite{goluskin&all2014}.  These runs are represented in Figs. 
\ref{fig:ma_v_eps}, \ref{fig:re_and_nu_v_ra}, \& \ref{fig:nrho_v_ra}
as hatched points, and while this phenomenon does not appear to greatly modify the
scaling of fluid properties measured in this work, these states warrant
further investigation.

@@

The efficiency of convection is quantified by the Nusselt number (Nu).  
Nu is well-defined in RBC
as the total flux normalized by the steady-state conductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficult to define, and we use
a modified version of a traditional stratified Nusselt number 
\cite{graham1975,hurlburt&all1984},
\begin{equation}
\text{Nu} \equiv \frac{\langle F_{\text{conv,z}} + F_{\text{cond,z}} - F_{\text{A}}\rangle}
{\langle F_{\text{cond,z}} - F_{\text{A}}\rangle} 
= 1 + \frac{\langle F_{\text{conv,z}}\rangle}{\langle F_{\text{cond,z}} - F_{\text{A}} \rangle}
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv,z}}$ and $F_{\text{cond,z}}$ are the 
z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
and $\langle \rangle$ are volume averages.  
$F_{\text{A}} \equiv -\langle\kappa\rangle \partial_z T_{\text{ad}}$ 
is the conductive flux of the proper corresponding adiabatic atmosphere.
For a compressible, ideal gas in hydrostatic equilibrium,
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ \cite{spiegel&veronis1960}.  
It is important to measure the evolved value of
$\langle \kappa \rangle = \langle \rho\chi \rangle$, which is nearly
$\kappa_0$ for small $\epsilon$ but changes appreciably for large
values of $\epsilon$.
In incompressible Boussinesq convection, where $\grad S = 0$ only when 
$\grad T = 0$, and where $\kappa$ is constant with depth and time,
this definition reduces to the traditionally defined
Nusselt number \cite{otero&all2002, johnston&doering2009}.
\begin{figure}[t!]
\includegraphics[width=\textwidth]{./figs/re_and_nu_v_Ra.png}
\caption{
Flow properties at high and low $\epsilon$. 
(a) Nu vs. Ra/Ra$_{\text{crit}}$.
Errors bars indicate the variance of Nu with depth;
large error bars indicate a poorly converged solution.
(b) Re vs. Ra/Ra$_{\text{crit}}$.
Re is measured at the midplane of the atmosphere.
Larger symbols with white dots designate 3D runs.
 \label{fig:re_and_nu_v_ra}
}
\end{figure}

@@

The rms Reynolds number (Re = $|\bm{u}|L_z/\nu$) and Peclet number
(Pe = Pr Re)
compare the importance of advection to diffusion in the evolved
convective state.  For Pr = 1, Pe = Re.  
Our choice of $\{\nu,\chi\}\propto \rho_0^{-1}$ drastically changes
the value of Re between the top and bottom of the atmosphere.  We report values of
Re at the midplane ($z=L_z/2$) of the atmosphere in
Fig. \ref{fig:re_and_nu_v_ra}b.  We find that Re
depends on largely Ra, but not $\epsilon$, except when the flow regime
changes.
In 2D Re $\appropto$ Ra$^{3/4}$ at low Ra.    When the 2D flows
become supersonic, 
Re $\appropto$ Ra$^{1/2}$, as expected
from (\ref{eqn:timescales}).
In our limited 3D runs,
Re $\appropto$ Ra$^{1/2}$, consistent with the supersonic results.
The heightened scaling
of Re in 2D follows the scaling of velocity (Ma) with Ra, as
seen in Fig. \ref{fig:ma_v_eps}, and reflects the presence of coherent
spinners, which do not exist in 3D.

@@

\bibliography{../biblio.bib}

>>>>

%\documentclass[iop]{emulateapj}
\documentclass[aps, pre, onecolumn, nofootinbib, notitlepage, groupedaddress, amsfonts, amssymb, amsmath, longbibliography]{revtex4-1}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}
\usepackage{bm}
\usepackage{natbib}
\usepackage{longtable}
\LTcapwidth=0.87\textwidth

@@

\newcommand\mnras{{MNRAS}}%

\begin{document}
\author{Evan H. Anders}
\affiliation{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affiliation{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}
\author{Benjamin P. Brown}
\affiliation{Dept. Astrophysical \& Planetary Sciences, University of Colorado -- Boulder, Boulder, CO 80309, USA}
\affiliation{Laboratory for Atmospheric and Space Physics, Boulder, CO 80303, USA}
\title{Convective heat transport in stratified atmospheres at low and high Mach number}

\begin{abstract}
We study fully compressible convection in the context of 
plane-parallel, polytropically stratified atmospheres. 
We perform a suite of 2D and 3D simulations in which we vary the initial
superadiabaticity ($\epsilon$) and the Rayleigh number (Ra) while fixing the
initial density stratification, aspect
ratio, and Prandtl number.
The evolved heat transport, 
quantified by the Nusselt number (Nu),
follows scaling relationships similar to those found in the well-studied, 
incompressible Rayleigh-B\'{e}nard problem.  This scaling holds up in both 2D and 3D and is not
appreciably affected by the magnitude of $\epsilon$.
\end{abstract}
\maketitle

\section{Introduction}
\label{sec:intro}
Convection transports energy in stellar and planetary atmospheres
where flows are compressible and
feel the atmospheric stratification.  This stratification 
is significant in regions such as
the convective envelope of the Sun, which spans 14 density scale heights.
In the bulk of these systems, particularly in the deep interior,
flows are at very low Mach number (Ma).  Unfortunately,
numerical constraints have restricted most studies of 
compressible convection to high Ma.
These prior studies \cite{graham1975, chan&all1982,
hurlburt&all1984, cattaneo&all1990, brummell&all1996,
brandenburg&all2005} have provided insight into the nature of
convection in the low temperature,
high Ma region near the Sun's surface. Few fundamental
properties of low Ma compressible convection, such as the scaling of
convective heat transport, are known.

@@

\section{Experiment} 
\label{sec:experiment}
We examine a monatomic ideal gas with an adiabatic index of
$\gamma = 5/3$ whose equation of state is $P = R\rho T$. This is consistent with the approach used in earlier work 
\cite{graham1975, chan&all1982, brandenburg&all2005,
hurlburt&all1984, cattaneo&all1990, cattaneo&all1991, brummell&all1996} 
and is the simplest stratified extension of RBC.
The atmospheres studied here are initially polytropically stratified,
\begin{equation}
\begin{split}
\rho_0(z) &= \rho_{t}(1 + L_z - z)^m, \\
T_0(z)    &= T_{t}(1 + L_z - z),
\label{eqn:polytrope}
\end{split}
\end{equation}
where $m$ is the polytropic index and $L_z$ is the depth of the atmosphere.
The polytropic
index is set by the superadiabatic excess, $\epsilon = m_{ad} - m$, where
$m_{ad} = (\gamma - 1)^{-1}$ is the adiabatic value of $m$.
The height coordinate, $z$, increases upwards in the range $[0, L_z]$.
Subscript 0 indicates initial conditions and subscript $t$ indicates values
at $z = L_z$.   Stratified atmospheres have a fourth non-dimensional parameter,
the number of density scale heights, $n_{\rho} = \ln\left[\rho_0(z=0)/\rho_t\right]$.  We
specify the depth of the atmosphere, $L_z = e^{n_{\rho}/m} - 1$, by choosing
the initial value of $n_{\rho}$.
Throughout this work we set $n_{\rho} = 3$.    Satisfying hydrostatic
equilibrium sets the value of gravity, $g = RT_t (m + 1)$, which is
constant with depth.  We study atmospheres with aspect
ratios of 4 where both the $x$ and $y$ coordinates have the range $[0, 4L_z]$.
In our 2D cases, we only consider $x$ and $z$.

@@

At fixed $n_\rho$, convective dynamics are 
controlled by $\epsilon$ as well as the atmospheric diffusivities.
At a fixed value of
$\epsilon$, the diffusivities are set by the
Rayleigh number (Ra) and the Prandtl number (Pr),
\begin{equation}
\text{Ra}_{t} = \frac{g L_z^3 (\Delta S_0 / c_P)}{\nu_t\chi_t},
\qquad
\text{Pr} = \frac{\nu}{\chi},
\end{equation}
where $\Delta S_0 = \epsilon\ln (1 + L_z) = \epsilon n_\rho / m$ 
is the initial specific entropy difference across the domain,
$\chi$ is the thermal diffusivity, 
and $\nu$ is the kinematic viscosity.
Throughout this work we specify
that Pr $= 1$ and is depth invariant.
The initial thermal
conductivity, $\kappa_0 = \chi \rho_0$, is
constant with depth, such that (\ref{eqn:polytrope}) is in
thermal equilibrium ($\grad\cdot[\kappa_0\grad T_0] = 0$).
By these
choices, $\nu(z) \equiv \chi(z) \equiv \chi_t / \rho_0$.
This formulation 
sets Ra at the bottom of the domain greater than
Ra$_t$ by a factor of $e^{2n_\rho}$. Henceforth
when we specify Ra we are referring to Ra$_t$.  
The full values of $\kappa = \rho\chi$ and 
$\mu = \rho\nu$ (the dynamic viscosity) change as the density 
profile evolves.  
The diffusivities scale as
$\chi_t, \nu_t \propto \sqrt{g L_z^3 (\Delta S_0 / c_P) / \text{Ra}_t}$.
Defining the thermal diffusion timescale as $t_\chi \equiv \tau_IL_z^2 / \chi$, the
ratio of $t_\chi$ to the buoyancy time is
\begin{equation}
\frac{t_\chi}{t_b} =\text{Ra}_t^{1/2}.
\label{eqn:timescales}
\end{equation}
We carry out two experiments in this study. In the first,
we fix $\epsilon$ and increase Ra, thus increasing the ratio in
(\ref{eqn:timescales}). In the second, we fix Ra and vary $\epsilon$,
scaling the dynamical timescales $(t_b, t_\chi)$ as $\epsilon^{-1/2}$
relative to the speed of sound; we see this reflected in the evolved Mach
number scaling (Fig. \ref{fig:ma_v_eps}).

We use $\ln \rho$ and $T$ as our thermodynamic variables and solve the Fully 
Compressible Navier-Stokes equations,
\begin{align}
&\begin{aligned}
&\frac{\partial \ln\rho}{\partial t} + \grad\cdot\bm{u} 
    = -\bm{u}\cdot\grad\ln\rho,
	\label{eqn:continuity_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial\bm{u}}{\partial t} + \grad T - 
&\nu\grad\cdot\lilstressT - \lilstressT\cdot\grad\nu =
-\bm{u}\cdot\grad\bm{u} - T\grad\ln\rho + \bm{g} + 
\nu\lilstressT\cdot\grad\ln\rho,
\label{eqn:momentum_eqn}
\end{aligned}\\
&\begin{aligned}
\frac{\partial T}{\partial t} -\frac{1}{c_V}\left(\right.\chi&\left.
    \grad^2 T + \grad T\cdot\grad\chi\right) =
	-\bm{u}\cdot\grad T - (\gamma-1)T\grad\cdot{\bm{u}}
	+ \frac{1}{c_V}\left(\chi\grad T \cdot\grad\ln\rho +
	\nu\left[\lilstressT\cdot\nabla\right]\cdot\bm{u}\right), 
	\label{eqn:energy_eqn}
\end{aligned}
\end{align}
with the viscous stress tensor given by
\begin{equation}
\sigma_{ij} \equiv \left(\frac{\partial u_i}{\partial x_j} + 
\frac{\partial u_j}{\partial x_i} - \frac{2}{3}\delta_{ij}\grad\cdot\bm{u}\right),
	\label{eqn:stress_tensor}
\end{equation}
where $\delta_{ij}$ is the Kronecker delta. Taking an inner product of
(\ref{eqn:momentum_eqn}) with $\rho\bm{u}$ and adding it to 
$\rho c_V\times$(\ref{eqn:energy_eqn}) reveals the full energy equation,
\begin{equation}
\frac{\partial}{\partial t}\left(\rho\left[\frac{|\bm{u}|^2}{2} + c_V T + \phi\right]\right) +
\Div{\bm{F}_{\text{conv}} + \bm{F}_{\text{cond}}} = 0,
	\label{eqn:energy_eqn_full}
\end{equation}
where
$
\bm{F}_{\text{conv}} \equiv \bm{F}_{\text{enth}} + \bm{F}_{\text{KE}} + \bm{F}_{\text{PE}} + \bm{F}_{\text{visc}}
$
is the convective flux and $\bm{F}_{\text{cond}} = -\kappa \grad T$
is the conductive flux.
The individual contributions to $\bm{F}_{\text{conv}}$ are the enthalpy flux, 
$\bm{F}_{\text{enth}} \equiv \rho\bm{u}(c_V T + P/\rho)$;
the kinetic energy flux, 
$\bm{F}_{\text{KE}} \equiv \rho|\bm{u}|^2\bm{u}/2$;
the potential energy flux,
$\bm{F}_{\text{PE}} \equiv \rho\bm{u}\phi$ (with $\phi \equiv -gz$);
and the viscous flux, 
$\bm{F}_{\text{visc}} \equiv -\rho\nu\bm{u}\cdot\lilstressT$.
Understanding how each of these fluxes interact  
is crucial in characterizing convective heat transport.

We utilize the 
Dedalus\footnote{\url{http://dedalus-project.org/}} 
pseudospectral framework \cite{burns&all2016} to time-evolve  
(\ref{eqn:continuity_eqn})-(\ref{eqn:energy_eqn}) 
using an implicit-explicit (IMEX), third-order, four-step 
Runge-Kutta timestepping scheme RK443 \cite{ascher&all1997}.  
Thermodynamic variables are decomposed such that $T = T_0 + T_1$ and
$\ln\rho = (\ln\rho)_0 + (\ln\rho)_1$, 
and the velocity is $\bm{u} = w\bm{\hat{z}} + u\bm{\hat{x}} + v\bm{\hat{y}}$.
In our 2D runs, $v = 0$.
Subscript 0 variables, set by (\ref{eqn:polytrope}), 
have no time derivative and vary only in $z$.
Variables are time-evolved on a dealiased Chebyshev (vertical)
and Fourier (horizontal, periodic) domain in which the
physical grid dimensions are 3/2 the size of the coefficient grid.  
Domain sizes range from
64x256 coefficients at the lowest values of 
Ra to 1024x4096 coefficients at Ra $> 10^{7}$ in 2D,
and from 64x128$^2$ to  256x512$^2$ in 3D. 
By using IMEX timestepping, we implicitly step the 
stiff linear acoustic wave contribution and are able to
efficiently study flows at high ($\sim 1$) 
and low ($\sim 10^{-4}$) Ma.  Our equations take the form
of the FC equations in \cite{lecoanet&all2014}, extended to include
$\nu$ and $\chi$ which vary with depth, and we follow the approach there.
This IMEX approach has been successfully 
tested against a nonlinear benchmark  of the compressible 
Kelvin-Helmholtz instability \cite{Lecoanet_et_al_2016_KH}.

@@

\begin{figure}[t]
\includegraphics[width=\textwidth]{./figs/ma_v_Ra.png}
\caption{The mean adiabatic Mach number of long-time-averaged profiles
is shown.  Error bars show the full range of Ma over the depth of the
atmosphere.
(a) Ma, at various values of Ra, 
is plotted as a function of $\epsilon$.
(b) Ma, at various values of $\epsilon$, 
is plotted as a function of Ra/Ra$_{\text{crit}}$.
Larger symbols with inlaid circles designate 3D runs.
\label{fig:ma_v_eps} }
\end{figure}

@@

We measure the adiabatic Mach number (Ma = $|\bm{u}|/\sqrt{\gamma T}$),
and find that it is a strong function of 
$\epsilon$ and a weak function of Ra.  
In our 2D runs, when Ma $< 1$, we observe a scaling law of
Ma(Ra$, \epsilon) \appropto \epsilon^{1/2}$Ra$^{1/4}$.
This relation breaks down as the mean
Ma approaches 1 (Fig. \ref{fig:ma_v_eps}).  This transition
occurs near Ra/Ra$_\text{crit} \approx \{10^{2}, 10^{3}\}$ for $\epsilon = \{1, 0.5\}$.
In our limited 3D runs, Ma appears to be a function of $\epsilon$ alone, with
Ma $\appropto \epsilon^{1/2}$, so at high Ra, Ma$_{\text{3D}} < \text{Ma}_{\text{2D}}$.
We conjecture that the scaling of Ma with Ra in the 2D runs 
is due to the formation of coherent high-velocity ``spinners,'' which form
between upflow and downflow lanes.  These structures, which are reminiscent of flywheel
modes in RBC, do not appear in our 3D runs at these parameters
\cite{jones&all1976, brummell&all2002}.
Simulations in the range of Ra/Ra$_{\text{crit}} > 10^3$ at $\epsilon = 10^{-4}$
exhibited ``windy'' states of convection, in which a large-scale horizontal
shearing flow replaced the more standard upflow/downflow morphology of
convection.  Similar states have been studied in
RBC \cite{goluskin&all2014}.  These runs are represented in Figs. 
\ref{fig:ma_v_eps}, \ref{fig:re_and_nu_v_ra}, \& \ref{fig:nrho_v_ra}
as hatched points, and while this phenomenon does not appear to greatly modify the
scaling of fluid properties measured in this work, these states warrant
further investigation.

@@

The efficiency of convection is quantified by the Nusselt number (Nu).  
Nu is well-defined in RBC
as the total flux normalized by the steady-state conductive flux 
\cite{johnston&doering2009, otero&all2002}.
In stratified convection Nu is more difficult to define, and we use
a modified version of a traditional stratified Nusselt number 
\cite{graham1975,hurlburt&all1984},
\begin{equation}
\text{Nu} \equiv \frac{\langle F_{\text{conv,z}} + F_{\text{cond,z}} - F_{\text{A}}\rangle}
{\langle F_{\text{cond,z}} - F_{\text{A}}\rangle} 
= 1 + \frac{\langle F_{\text{conv,z}}\rangle}{\langle F_{\text{cond,z}} - F_{\text{A}} \rangle}
\label{eqn:nusselt}
\end{equation}
where $F_{\text{conv,z}}$ and $F_{\text{cond,z}}$ are the 
z-components of $\bm{F}_{\text{conv}}$ and $\bm{F}_{\text{cond}}$,
and $\langle \rangle$ are volume averages.  
$F_{\text{A}} \equiv -\langle\kappa\rangle \partial_z T_{\text{ad}}$ 
is the conductive flux of the proper corresponding adiabatic atmosphere.
For a compressible, ideal gas in hydrostatic equilibrium,
$\partial_z T_{\text{ad}} \equiv - g / c_{P}$ \cite{spiegel&veronis1960}.  
It is important to measure the evolved value of
$\langle \kappa \rangle = \langle \rho\chi \rangle$, which is nearly
$\kappa_0$ for small $\epsilon$ but changes appreciably for large
values of $\epsilon$.
In incompressible Boussinesq convection, where $\grad S = 0$ only when 
$\grad T = 0$, and where $\kappa$ is constant with depth and time,
this definition reduces to the traditionally defined
Nusselt number \cite{otero&all2002, johnston&doering2009}.
\begin{figure}[t!]
\includegraphics[width=\textwidth]{./figs/re_and_nu_v_Ra.png}
\caption{
Flow properties at high and low $\epsilon$. 
(a) Nu vs. Ra/Ra$_{\text{crit}}$.
Errors bars indicate the variance of Nu with depth;
large error bars indicate a poorly converged solution.
(b) Re vs. Ra/Ra$_{\text{crit}}$.
Re is measured at the midplane of the atmosphere.
Larger symbols with inlaid circles designate 3D runs.
 \label{fig:re_and_nu_v_ra}
}
\end{figure}

@@

The rms Reynolds number (Re = $|\bm{u}|L_z/\nu$) and Peclet number
(Pe = Pr Re)
compare the importance of advection to diffusion in the evolved
convective state.  For Pr = 1, Pe = Re.  
Our choice of $\{\nu,\chi\}\propto \rho_0^{-1}$ drastically changes
the value of Re between the top and bottom of the atmosphere.  We report values of
Re at the midplane ($z=L_z/2$) of the atmosphere in
Fig. \ref{fig:re_and_nu_v_ra}b.  We find that Re
depends largely on Ra, but not $\epsilon$, except when the flow regime
changes.
In 2D Re $\appropto$ Ra$^{3/4}$ at low Ra.    When the 2D flows
become supersonic, 
Re $\appropto$ Ra$^{1/2}$, as expected
from (\ref{eqn:timescales}).
In our limited 3D runs,
Re $\appropto$ Ra$^{1/2}$, consistent with the supersonic results.
The heightened scaling
of Re in 2D follows the scaling of velocity (Ma) with Ra, as
seen in Fig. \ref{fig:ma_v_eps}, and reflects the presence of coherent
spinners, which do not exist in 3D.

@@
